---
title: "SQL"
output: html_document
---

## Filtering Columns and Rows

With SQL, you can filter columns and rows by using SELECT and WHERE. Let's look at an exmaple using the Lahman database. So we should first load that database, along with the package sqldf.

```{r message=FALSE, warning=FALSE}
library(Lahman)
library(sqldf)
```

Suppose now I would like to see the homerun totals for the 1927 Yankees. I could just as easily writing the following: 
```{r}
query<-"select playerID, yearID, teamID, HR FROM Batting
WHERE teamID='NYA' and yearID=1927"
sqldf(query)
```

Let's look at another example. Let's say I want all instances where the Yankees hit 40 homeruns or more. That would look like this:
```{r}
query<-"select playerID, yearID, teamID, HR FROM Batting
WHERE teamID='NYA' and HR>=40"
sqldf(query)
```

Let's look at example with more than one constraint. Suppose we want more than 40 homeruns, but less than 60 strikeouts:
```{r} 
query<-"select playerID, yearID, teamID, HR, SO FROM Batting
WHERE HR>40 and SO<60"
sqldf(query)
```

For a final test, let's add two constraints to the same category. Let's discriminate Phillies with more than 30 Homeruns during the 1970s.
```{r}
query<-"select playerID, yearID, teamID, HR FROM Batting
WHERE teamID='PHI' AND HR>30 and yearID>1969 and yearID<1980"
sqldf(query)
```

## ORDER BY
Here's an example of ordering, where we order the rows by a different constraint other than chronological.
```{r}
query<- "SELECT playerID, teamID, yearID, HR FROM Batting
WHERE HR>50
ORDER BY HR DESC"
sqldf(query)
```

Here's an exmaple with two constraints, strikeouts and the number of at bats:
```{r}
query<-"select playerID, yearID, teamID, AB, SO FROM Batting
WHERE SO<10 and AB>=400
ORDER BY SO"
sqldf(query)
```


## AGGREGATION
Aggregation is a way to combine multiple rows into a single data output. This includes averaging, sum, mins or maxes of a data range.

Our first example is to find the average homes Babe Ruth had throughout his career.
```{r}
query<-"select playerID, sum(HR) from Batting where playerID='ruthba01'
group by playerID"
sqldf(query)
```

Our second example to is find all players who had over 600 homeruns throughout their career. That looks like this.
```{r}
query<-"select playerID, sum(HR) from Batting 
group by playerID 
having sum(HR)>=600
order by sum(HR) desc"
sqldf(query)
```


Our final example is to rank the players by their average homeruns per season, which combines several intuitive code forms.
```{r}
query<-"select PlayerID, avg(hr) FROM Batting
GROUP BY playerID
HAVING avg(HR)>30
ORDER BY AVG(HR)>30 DESC"
sqldf(query)
```


## JOINS

JOINS are a way to combine data from difference data sets, IE Batting and Master. To do such a thing, for example, replacing ruthba01 with Babe Ruth, looks like this.
```{r}
query<-"select teamID, nameFirst, nameLast, yearID, HR
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
WHERE Batting.playerID='ruthba01'"
sqldf(query)
```

Our second example here is to display all players who hit more than 50 homeruns, but to replace their playerID with their real first and last names.
```{r}
query<-"select teamID, HR, nameFirst, nameLast, yearID
From Batting INNER JOIN Master
ON Batting.playerID=Master.PlayerID
WHERE HR>50
ORDER BY HR"
sqldf(query)
```


Our third example is to replace babe Ruth's teamIDs with the actual team name, this time taking it from the "Teams" column.
```{r}

query<-"Select playerID, name, Batting.yearID, Batting.HR
FROM Batting INNER JOIN Teams
ON Batting.teamID=Teams.teamID and Batting.yearID=Teams.yearID
WHERE playerID='ruthba01'"
sqldf(query)
```

Here's another example where we display both Babe Ruth's actual name and the actual team he played for in a given year.
```{r}
query<-"SELECT  nameFirst, nameLast, name, Batting.HR, Batting.yearID 
FROM Batting INNER JOIN Teams INNER JOIN Master
ON Batting.teamID=Teams.teamID and Batting.yearID=Teams.yearID and batting.playerID=Master.PlayerID
where batting.playerID='ruthba01'"
sqldf(query)

```

##Additional Examples
This example gives all players with more than 600 career homeruns, this time giving us their real names.
```{r}
query<-"Select namefirst, nameLast, sum(Batting.HR)
from Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
GROUP BY Batting.playerID
Having sum(Batting.HR)>600
ORDER BY sum(HR) DESC"
sqldf(query)
```


sqldf(query)
```{r}
query<-"Select salary, nameFirst, nameLast
FROM Batting INNER JOIN Master INNER Join Salaries
on Batting.playerID=Master.PlayerID=Salaries.playerID
GROUP by Batting.playerID
Having avg(salary)>1000000
ORDER BY avg(salary) DESC"
sqldf(query)
```
